---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

### TO-DO:

1.  Re asses handling of missing hours -\> should be treated as 0? Munyo does not do that, but it seems useful
2.  Why are points in rdplot() never above one another?

### The dataset stems from the city of New York <insert link>.

```{r}
library(pacman)
p_load(tidyverse)
nypd_complaints <- read_csv("NYPD_Complaint_Data_Historic.csv")
```

### Filtering the dataset to only include robbery and burglary (this is not quite equivalent to robbery and theft, but close enough).

```{r}
# Create the new dataframe with only relevant rows and label them
robburg <- nypd_complaints %>%
  filter(OFNS_DESC %in% c(
    # Theft-related offense descriptions
    "GRAND LARCENY",
    "PETIT LARCENY",
    "GRAND LARCENY OF MOTOR VEHICLE",
    "PETIT LARCENY OF MOTOR VEHICLE",
    "THEFT-FRAUD",
    "THEFT OF SERVICES",
    "OTHER OFFENSES RELATED TO THEF",
    "POSSESSION OF STOLEN PROPERTY",
    "FRAUDS",
    "FORGERY",
    "FRAUDULENT ACCOSTING",
    "JOSTLING",
    
    # Robbery-specific (only one)
    "ROBBERY"
  )) %>%
  mutate(THEFT_OR_ROBBERY = case_when(
    OFNS_DESC == "ROBBERY" ~ "ROBBERY",
    TRUE ~ "THEFT"
  ))
head(robburg)
```

### Now a little bit of restructuring!

```{r}
# CHECK if anything of this is relevant later on!!!

robburg <- robburg %>%
  mutate(
    # Convert CMPLNT_FR_DT (time of incoming 911 call) to Date
    date_parsed = mdy(CMPLNT_FR_DT),
    # Combine date and time into a POSIXct datetime
    startdt = as.POSIXct(date_parsed + CMPLNT_FR_TM, origin = "1970-01-01", tz = "America/New_York")
  )

# ---------------------

robburg <- robburg %>%
  mutate(date_parsed = mdy(CMPLNT_FR_DT))

# Extract year from CMPLNT_FR_DT
robburg <- robburg %>%
  mutate(year_parsed = year(date_parsed))

# Some years in CMPLNT_FR_DT are incorrect, so I fix it using the year from CMPLNT_TO_DT. I only do so when the year from CMPLNT_FR_DT exceeds the theoretical timeframe. There is still a possibility that some of the years are not correct.
robburg <- robburg %>%
  mutate(
    alt_date = mdy(CMPLNT_TO_DT),
    alt_year = year(alt_date),
    corrected_year = case_when(
      is.na(date_parsed) ~ NA_integer_,
      year_parsed < 2006 | year_parsed > 2025 ~ alt_year,
      TRUE ~ year_parsed
    )
  )

# Because some faulty rows do not include CMPLNT_FR_DT, in the last step, I only keep rows with valid years.
robburg <- robburg %>%
  filter(!is.na(corrected_year) & corrected_year >= 2006 & corrected_year <= 2025)

# year is being replaced with the correct year
robburg <- robburg %>%
  mutate(year = corrected_year) %>%
  select(-year_parsed, -alt_year, -corrected_year)


```

### Now I include and structure the data describing DST dates (and time... which is always 2am)

```{r}
# import dates from self-made csv file. Data from some website.
dst_dates <- read_delim("dst_dates.csv", 
                        delim = ";", escape_double = FALSE, trim_ws = TRUE)
# remove empty row
dst_dates <- dst_dates[-20,]

dst_spring <- dst_dates$dst_start_date

# Create a vector with all dates a week before and after the DST change in spring
dst_period <- map(dst_spring, ~ seq(.x, .x + 6, by = "day")) |> unlist() |> as.Date(origin = "1970-01-01")

# Some column-creating
robburg$month <- month(as.Date(robburg$startdt))
robburg$day <- day(as.Date(robburg$startdt))
robburg$date <- as.Date(sprintf("%04d-%02d-%02d",
                                as.integer(robburg$year),
                                as.integer(robburg$month),
                                as.integer(robburg$day)))
robburg$hour <- hour(robburg$CMPLNT_FR_TM)
robburg$weekday <- wday(robburg$date)

# Assign treatment
robburg <- robburg %>%
  mutate(
    DST = case_when(
      date %in% dst_spring & hour < 2 ~ 0,
      date %in% dst_period ~ 1,
      TRUE ~ 0
    )
  )
```

### Darkness dummy

```{r}
p_load(suncalc)

# Get unique dates. Latitude and Longitude are static for the centre of NYC
nyc_dates <- robburg %>%
  filter(!is.na(date)) %>%
  distinct(date) %>%
  mutate(
    lat = 40.7128,
    lon = -74.0060
  )

# Get sunrise/sunset times for NYC center
nyc_suntimes <- getSunlightTimes(
  data = nyc_dates,
  keep = c("sunrise", "sunset"),
  tz = "America/New_York"
)

robburg <- robburg %>%
  left_join(nyc_suntimes, by = "date")

# Construct datetime
robburg <- robburg %>%
  mutate(
    hour_time = as.POSIXct(paste(format(date), sprintf("%02d:00:00", hour)),
                           format = "%Y-%m-%d %H:%M:%S",
                           tz = "America/New_York")
  )


# Create darkness dummy
robburg <- robburg %>%
  mutate(
    darkness = if_else(hour_time < sunrise | hour_time >= sunset, 1, 0)
  )
```

### Create crime_hourly

The dependent variable is crime per hour, so I need to aggregate the data from above.

```{r}
crime_hourly <- robburg %>% group_by(date, hour, year, month, weekday, DST, darkness) %>% summarise(crime_count = n(), .groups = "drop")
head(crime_hourly)
```

```{r}
crime_hourly <- robburg %>%
  group_by(date, hour, year, month, weekday, DST, darkness) %>%
  summarise(
    crime_count = n(),
    robbery = sum(THEFT_OR_ROBBERY == "ROBBERY", na.rm = TRUE),
    theft = sum(THEFT_OR_ROBBERY == "THEFT", na.rm = TRUE),
    .groups = "drop"
  )
```

### Add running variable

The running variable is the distance from the DST change in spring (in hours).

```{r}
dst_spring.df <- tibble(year = year(as.Date(dst_spring)))

dst_spring.df$dst_date <- as.Date(dst_spring)

crime_hourly <- crime_hourly %>%
  left_join(dst_spring.df, by = "year")


crime_hourly <- crime_hourly %>%
  mutate(
    # Construct the datetime correctly using `hour` column
    datetime = as.POSIXct(paste(date, sprintf("%02d", .data$hour)), format = "%Y-%m-%d %H", tz = "UTC"),
    
    # Assume DST change happens at 2 AM (can be adjusted if needed)
    dst_change_datetime = as.POSIXct(paste(dst_date, "02"), format = "%Y-%m-%d %H", tz = "UTC"),
    
    # Calculate running variable (in hours from DST change)
    running = as.numeric(difftime(datetime, dst_change_datetime, units = "hours"))
  )
```

Removing rows that do not have a value for darkness. 14 rows are dropped.

```{r}
crime_hourly <- crime_hourly %>% filter(is.na(darkness) == FALSE)
```

```{r}
crime_hourly.c <- crime_hourly %>% filter(is.na(darkness) == FALSE) # drops 14 rows 
```

Create dummy variables for the covariates, because rdrobust cannot handle covariates as factors directly, so I need to create a matrix with 0/1 coded values.

```{r}
# Create dummy variables for categorical covariates
x_dummies <- model.matrix(~ factor(year) + factor(month) + factor(hour) + factor(weekday) + factor(darkness) - 1, data = crime_hourly.c)
```

### Modeling RDDs

```{r}
p_load(rdrobust)
summary(with(crime_hourly.c, rdrobust(crime_count, running, c=0, p=2, h=72)))
summary(with(crime_hourly.c, rdrobust(crime_count, running, c=0, p=2, h=72, covs=x_dummies)))
```

```{r}
summary(with(crime_hourly.c, rdrobust(crime_count, running, c=0, p=2, h=96)))
summary(with(crime_hourly.c, rdrobust(crime_count, running, c=0, p=2, h=96, covs=x_dummies)))
```

```{r}
summary(with(crime_hourly.c, rdrobust(crime_count, running, c=0, p=2, h=120)))
summary(with(crime_hourly.c, rdrobust(crime_count, running, c=0, p=2, h=120, covs=x_dummies)))
```

### Plotting RDDs

```{r}
with(crime_hourly.c,
     rdplot(
        crime_count,
        running,
        c=0,
        p=2,
        h=120,
        x.lim = c(-168, 168),
       title = "Effect of DST Change on Hourly Crime Rates",
       x.label = "Distance from Spring DST Change in Hours",
       y.label = "Number of Crimes per Hour"
))
```

```{r}
with(crime_hourly.c,
     rdplot(
       crime_count,
       running,
       c=0,
       p=2,
       h=96,
       x.lim = c(-120, 120),
       title = "Effect of DST Change on Hourly Crime Rates",
       x.label = "Distance from Spring DST Change in Hours",
       y.label = "Number of Crimes per Hour"
     ))
```

```{r}
with(crime_hourly.c,
     rdplot(
       crime_count,
       running,
       c=0,
       p=2,
       h=72,
       x.lim = c(-96, 96),
       nbins = c(100000, 100000),
       title = "Effect of DST Change on Hourly Crime Rates",
       x.label = "Distance from Spring DST Change in Hours",
       y.label = "Number of Crimes per Hour"
     ))
```

```{r}
library(ggplot2)
library(dplyr)

# Filter data to +/- 96 hours
crime_hourly.filtered <- crime_hourly.c %>%
  filter(running >= -96, running <= 96)

# Split data
left <- crime_hourly.filtered %>% filter(running < 0)
right <- crime_hourly.filtered %>% filter(running >= 0)

# Modeling
model_left <- lm(crime_count ~ running, data = left)
model_right <- lm(crime_count ~ running, data = right)

# Add predicted values for plotting lines
left <- left %>% mutate(predicted = predict(model_left))
right <- right %>% mutate(predicted = predict(model_right))

# Plot
ggplot(crime_hourly.filtered, aes(x = running, y = crime_count)) +
  geom_point(alpha = 0.2, color = "pink") +
  geom_line(data = left, aes(x = running, y = predicted), color = "darkblue", size = 1) +
  geom_line(data = right, aes(x = running, y = predicted), color = "darkblue", size = 1) +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(
    title = "Effect of DST Change on Hourly Crime Rates",
    x = "Distance from Spring DST Change in Hours",
    y = "Number of Crimes per Hour"
  ) +
  theme_light()


```

### 
